Test mr.ash2s on ukbiobank data, chr17-22

### 2020-09-03 analysis/index.Rmd
susie prior

  * [Using UKbiobank data (chr17 to 22), fusion weights Adipose subcutaneous](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.html)

@@

@@

  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17to22-gtex.adipose.html) 
  * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)

>>>>

@@

# Test parameters for susie
 * [Using UKbiobank data (chr17:22), fusion weights Adipose, test prior's effect for susie](simulation-multi-ukbchr17to22-gtex.adipose_susieprior.html)
    * [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-susieP-ukbchr17to22-gtex.adipose.html)

@@

  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17to22-gtex.adipose.html) 
  * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)
      *   [Visualization](https://users.rcc.uchicago.edu/~simingz/simulation-mr.ash2s-ukbchr17to22-gtex.adipose.html)

## mr.ash + SUSIE iterative
*  [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-susieI-ukbchr17to22-gtex.adipose.html)

### 2020-09-03 analysis/simulation-multi-ukbchr17to22-gtex.adipose_susieprior.Rmd
susie prior

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

@@

## Truth

## Using uniform prior:

```{r cali 0, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

## Use overestimated prior

## Use priors close to truth

```{r cali 2, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

## Use under estimated priors

* susie results: 
```{r cali 3, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

We run 100 simulations and run susie using different priors and `L= 1`. We apply susie for all regions.

## Truth

@@

## Using uniform prior:

@@

## Use overestimated prior

@@

## Use priors close to truth

* The prior used is: 
```{r  par2 all}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

```{r cali 2 all, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

## Use under estimated priors

>>>>

## PIP calibration: filter regions

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-1-fixprior_rpip0.5/"
```

We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with mr.ash2 regional PIP > 0.5.

### Truth

### Using uniform prior:

```{r cali 0, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
```

### Use overestimated prior

### Use priors close to truth

```{r cali 2, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

### Use under estimated priors

* susie results: 
```{r cali 3, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## PIP calibration: all regions

```{r output a dir}
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

@@

### Truth

* The true parameters we used to simulate data: 
```{r  true a par}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[,2, drop = F])
```

### Using uniform prior:

```{r cali 0 a, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

### Use overestimated prior

* The prior used is: 
```{r  par1 a}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_1.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 1 a, fig.height=6, fig.width=14}
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

### Use priors close to truth

* The prior used is: 
```{r  par2 a}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

```{r cali 2 a, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

### Use under estimated priors

* The prior used is: 
```{r  par3 a}
par <- read.table(paste0(susiedir, "20200721-1-fixedprior_3.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

* susie results: 
```{r cali 3 a, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-1-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

## PIP calibration: filter regions

```{r output dir-2}
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/20200721-3-fixprior_causal/"
```

We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for regions with at least one causal signal.

### Truth

* The true parameters we used to simulate data: 
```{r  true par-2}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
t(par[,2, drop = F])
```

### Using uniform prior:

```{r cali 0-2, fig.height=6, fig.width=6}
pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior1.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
cp_plot(res$pip.null, res$ifcausal, main = "SUSIE.uniform PIP")
```

### Use overestimated prior

* The prior used is: 
```{r  par1-2}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_1.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 1-2, fig.height=6, fig.width=14}
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

### Use priors close to truth

* The prior used is: 
```{r  par2-2}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 2-2, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

### Use under estimated priors

* The prior used is: 
```{r  par3-2}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_3.txt"), header = T)
t(par[c(1,3),1, drop = F])
```

* susie results: 
```{r cali 3-2, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior3.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))
par(mfrow=c(1,2))
cp_plot(res$pip, res$ifcausal, main = "SUSIE.weighted PIP")
cp_plot(res$pip.w0, res$ifcausal, main = "SUSIE.weighted-w-null PIP")
```

## PIP calibration: all regions

```{r output dir all}
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

We run 100 simulations and run susie using different priors and `L= 1`. We only apply susie for all regions.

### Truth

@@

### Using uniform prior:

@@

### Use overestimated prior

@@

### Use priors close to truth

* The prior used is: 
```{r  par2 all}
par <- read.table(paste0(susiedir, "20200721-3-fixedprior_2.txt"), header = T)
t(par[c(1,3),1, drop = F])
```
* susie results: 
```{r cali 2 all, fig.height=6, fig.width=14}
pipfs <- Sys.glob(paste0(susiedir,"20200721-3-*.fixedprior2.L1.susieres.expr.txt"))
res <- do.call(rbind, lapply(pipfs, read.table, header = T))

@@

### Use under estimated priors

### 2020-09-03 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd
susie prior

@@

>>>>

* Susie run parameters: `L=5`. Update `null_weight` in each iteration. We initialize prior for gene as `0.05` and prior for SNPs as `2.5E-3`, we initialize `null_weight` as 0.

@@

# susieI genes - setting 6

Prior change for each iteration:
```{r iter6}
load(paste0(outputdir, tag, ".config5.susieIres.Rd"))
df <- cbind(prior.gene_rec, prior.SNP_rec)
rownames(df) <- paste("Iteration", 1:Niter)
colnames(df) <- c("Prior.gene", "Prior.SNP")
df
```

```{r param6}
mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config5.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config5.susieI.txt")

### 2020-09-03 analysis/simulation-susieP-ukbchr17to22-gtex.adipose.Rmd
susie prior

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})

source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

plot_pip <- function(dt){

  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  if ( "max_r2" %in% colnames(dt)) {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
                   mode = 'markers', color = ~ ifcausal, colors = pal,
                   text = ~ paste("Name: ", name, 
                                  "\nmax R^2: ", round(max_r2, digits = 2)))
                                  #,"\nMax pair: ", max_pair))
  } else {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
  }

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_beta <- function(dt){
  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')

  fig <- fig %>% layout(yaxis = list(title = "beta"))
  return(fig)
}

plot_p <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "p value"))
  return(fig)
}

plot_susie_pip <- function(dt) {
   pal <- c("salmon", "darkgreen")
   pal <- setNames(pal, c("Causal", "Non causal"))

   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres.gene, susieres.snp, ldres = NULL, chrom=NULL){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  if (!is.null(ldres)){
    ld <- read.table(ldres, header =T)
    a <- merge(a, ld, by = "name", all.x = T)}

  su <- read.table(susieres.gene, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

  p <- fread(gwasres.gene, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.gene <- merge(a, p, by = "name", all.x = T)

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  su <- read.table(susieres.snp, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

  p <- fread(gwasres.snp, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.snp <- merge(a, p, by = "name", all.x = T)

  if (!is.null(chrom)){
    dt.gene <- dt.gene[dt.gene$chr.x == chrom,]
    dt.snp <- dt.snp[dt.snp$chr.x == chrom,]
  }

  fig1 <- plot_pip(dt.gene)  
  fig2 <- plot_beta(dt.gene) 
  fig3 <- plot_p(dt.gene) 
  fig4 <- plot_susie_pip(dt.gene) 
  fig5 <- plot_pip(dt.snp) 
  fig6 <- plot_beta(dt.snp) 
  fig7 <- plot_p(dt.snp)
  fig8 <- plot_susie_pip(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, fig8, nrows=8, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 1100, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

```

@@

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
  gsusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.expr.txt")
  ssusie <- paste0(susiedir, tag, ".", tag2, ".L3.susieres.snp.txt")
  rsq <- paste0(outputdir, tag, ".LD.txt")
  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, gsusie, ssusie, rsq))
}
```

>>>>

@@

```{r source file}
source("analysis/visual_twas_plots.R")
```
